---
title: "HW4"
author: "Drew Andersen"
date: "5/18/2022"
output: html_document
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(here)
library(tidyverse)
library(kableExtra)
```

#Introduction
```{r}
damages <- read_csv(here("damages.csv"))
warming <- read_csv(here("warming.csv"))[,-1]

#changing names
damages <- damages %>% 
  rename(dam = damages,
         deg_c = warming) %>% 
  mutate(deg_c_sq = deg_c^2)

warming <- warming %>% 
  rename(CO2_base = warming_baseline,
         CO2_pulse = warming_pulse)
```

##Question 1
```{r}
#quad model with 0 intercept
quad_model <- lm(dam ~ 0 + deg_c + deg_c_sq, data = damages)

#function for quadratic fit 
quad_fun <- function(deg_c, deg_c_sq){
  dam <- quad_model$coefficients[1]*deg_c + quad_model$coefficients[2]*deg_c_sq
  return(dam)
}

damages$dam_fit = quad_fun(damages$deg_c, damages$deg_c_sq)
```

```{r}
ggplot(data = damages, aes(x = deg_c, y = dam)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) +
  xlab("Warming (˚C)") +
  ylab("Annual Total Damages ($)") +
  theme_minimal()
```
 
 **Figure 1:** The annual total damages from climate change warming in degree Celsius fitted with a quadratic function with a zero intercept. 
 
##Question 2
```{r}
#using function to calculate env damage from baseline, pulse, difference, and pulse per ton of CO2
warming$base_dam <- quad_fun(warming$CO2_base,(warming$CO2_base)^2)
warming$pulse_dam <- quad_fun(warming$CO2_pulse,(warming$CO2_pulse)^2)
warming$dam_diff <- warming$pulse_dam - warming$base_dam
warming$pulse_dam_per_ton_CO2 <- warming$dam_diff/(35E9)
```

```{r}
ggplot(data = warming, aes(x = year, y = base_dam)) +
  geom_point() +
  xlab("Year") +
  ylab("Annual Total Damages ($)")
```

**Figure 2:** Damages over time without the pulse.

```{r}
ggplot(data = warming, aes(x = year, y = pulse_dam)) +
  geom_point() +
  xlab("Year") +
  ylab("Annual Total Damages ($)") 
```

**Figure 3:** Damages over time with the pulse.

```{r}
ggplot(data = warming, aes(x = year, y = dam_diff)) +
  geom_point() +
  xlab("Year") +
  ylab("Annual Total Damages ($)") 
```

**Figure 4:** The difference in damages over time that arises from the pulse.

```{r}
ggplot(data = warming, aes(x = year, y = pulse_dam_per_ton_CO2)) +
  geom_point() +
  xlab("Year") +
  ylab("Annual Total Damages ($)") 
```

**Figure 5:** The difference in damages over time from the pulse per ton of CO2 (assuming that each ton of the pulse causes the same amount of damage).

##Question 3 & 4
```{r}
one_ton_dam = sum(warming$pulse_dam_per_ton_CO2)

discount_fun <- function(rate, time, fv){
  npv <- fv/(1+(rate/100))^time
  return(npv)
}

scc_df <- data_frame(rate = seq(0,10,0.2))
scc_df$scc <- discount_fun(scc_df$rate, 78, one_ton_dam)

#q4 stuff
ramsey_rate = (.01 + (2*.01)) * 100 #multiplied by 100 to be compatible with function
discount_ramsey = discount_fun(ramsey_rate, 78, one_ton_dam)

ggplot(data = scc_df, aes(x = rate, y = scc)) +
  geom_point() +
  geom_point(aes(x = ramsey_rate, y = discount_ramsey), colour = "red")
```

**Figure 6:** The social cost of carbon (SCC) at various discount rates. A higher discount rate leads to a lower SCC. The red dot represent the SCC using the Ramsey Rule (rate = `r ramsey_rate`%).

##Question 5
```{r}
rate_q5 = 2
#Policy A (BAU)


#Policy B (Strong Action)
warming$policy_B <- discount_fun(rate_q5, (warming$year - 2022), warming$base_dam)
adding = seq(warming$policy_B[1], warming$policy_B[28], 1)
epv_to_2050 = mapply(sum, c(warming$policy_B[1]:warming$policy_B[28]))
epv_post_2050 = warming$policy_B[30]*50 #50 year period
```

